Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

Cryptographic hash functions form the core of many protocols which
trust that these functions preserve certain basic
properties. Cryptographic hash functions see widespread
use for file integrity checks to verify long term storage of
data, as cache invalidation techniques, and as a building block in network
protocols such as Kerberos and TLS.  The classes of functions suitable
for these use cases necessarily have
strong guarantees about properties of its members.
An ideal cryptographic hash has five main features: determinism - the
same message always results in the same hash; performability -
computing a hash for a message should not be algorithmically
difficult; non-reversibility - it should be infeasible to generate a message
from its hash except by trying all possible messages; sensitivity to
initial conditions - small changes to a message should change every bit of
the hash value with probability $\frac{1}{2}$;
unique digests - it should be infeasible to find two different messages with
the same hash value.

@@

We seek to address the problem of evaluating the use cases for a collision,
to determine whether or not a collision affects a particular system.
The central thesis of this research is that logical cryptanalysis
allows us to formalize the structure inherent to the underlying hash
function such that new collisions can be found within reasonable time
on commodity hardware.  We assert that when such structure exists, it
is \textbf{imperitive to discontinue use of such hash functions as
  they are fundamentally untrustworthy for critical applications.}  An
example of such structure showing the unit distance neighborhood
(defined in Section \ref{Sec:Terminology}) of the known collisions
from Dobbertin and Kasselman is presented in Figure \ref{Fig:Graph-Intro}.
This paper contributes several techniques to logical cryptanalysis and uses these
techniques to derive metrics of the trustworthiness for hash functions. We claim
the following results are novel in the field:
\begin{itemize}
    \item We define the neighborhood of a class of collisions and show that
        it is frequently non-empty.
    \item We discuss techniques for finding useful collisions within a
        class using logical cryptanalysis.
    \item We propose additional techniques for building new classes of
        collisions and show that classes of collisions in MD4 can be found inductively.
    \item We propose five universal metrics for evaluating the utility of
        a class of collisions.
\end{itemize}
We propose that collision classes of high utility are closer to a second
preimage attack in the amount of flexibility they provide to an attacker,
whereas classes of low utility may not affect all systems which use a
particular hash function.

@@

We have demonstrated in this paper that our theories for logical
cryptanalysis hold, and lead to a 35 fold increase in collision
discovery when compared to the existing literature.  By employing our
presented distance heuristic and the lattice structure uncovered by
the $\precsim$ relation, we can evaluate the trustworthiness of a hash
function as the inverse of the flexibility of known collisions for
that hash function.  This allows us to measure the structural
weaknesses of a hash function and determine if extant collisions are sufficient
to render a hash function untrustworthy, and likewise to understand if
necessary conditions related to the underlying structure are met to
ensure trustworthiness.

>>>>

Cryptographic hash functions form the core of many protocols which
trust that these functions preserve certain basic
properties. Cryptographic hash functions see widespread
use as file integrity checks to verify long term storage of
data, as cache invalidation techniques, and as a building block in network
protocols such as Kerberos and TLS.  The classes of functions suitable
for these use cases necessarily have
strong guarantees about the properties of its members.
An ideal cryptographic hash has five main features: determinism - the
same message always results in the same hash; performability -
computing a hash for a message should not be algorithmically
difficult; non-reversibility - it should be infeasible to generate a message
from its hash except by trying all possible messages; sensitivity to
initial conditions - small changes to a message should change every bit of
the hash value with probability $\frac{1}{2}$;
unique digests - it should be infeasible to find two different messages with
the same hash value.

@@

We seek to address the problem of evaluating the use cases for a collision,
to determine whether or not a collision affects a particular system.
The central thesis of this research is that logical cryptanalysis
allows us to formalize the structure inherent to the underlying hash
function such that new collisions can be found within reasonable time
on commodity hardware.  We assert that when such structure exists, it
is \textbf{imperitive to discontinue use of such hash functions as
  they are fundamentally untrustworthy for critical applications.}  An
example of such structure showing the unit distance neighborhood
(defined in Section \ref{Sec:Terminology}) of the known collisions
from Dobbertin and Kasselman is presented in Figure \ref{Fig:Graph-Intro}.

This paper contributes several techniques to logical cryptanalysis and uses these
techniques to derive metrics of the trustworthiness for hash functions. We claim
the following results are novel in the field:
\begin{itemize}
    \item We define the neighborhood of a class of collisions and show that
        it is frequently non-empty.
    \item We discuss techniques for finding useful collisions within a
        class using logical cryptanalysis.
    \item We propose additional techniques for building new classes of
        collisions and show that classes of collisions in MD4 can be found inductively.
    \item We propose five universal metrics for evaluating the utility of
        a class of collisions.
\end{itemize}
We propose that collision classes of high utility are closer to a second
preimage attack in the amount of flexibility they provide to an attacker,
whereas classes of low utility may not affect all systems which use a
particular hash function.

@@

We have demonstrated in this paper that our theories for logical
cryptanalysis hold: namely the existance of useful distance
heuristics; the existance of a similarity operator, $\precsim$,
between families of collision classes; and the existance of a useful
heuristic structure for second preimage and chosen prefix attacks.
These theories lead to techniques which result in a 35 fold increase in collision
discovery when compared to the existing literature.  By employing our
presented distance heuristic and the lattice structure uncovered by
the $\precsim$ relation, we can evaluate the trustworthiness of a hash
function as the inverse of the flexibility of known collisions for
that hash function.  This allows us to measure the structural
weaknesses of a hash function and determine if extant collisions are sufficient
to render a hash function untrustworthy, and likewise to understand if
necessary conditions related to the underlying structure are met to
ensure trustworthiness.
